Grappa: Scalable Computing on Mass-Market Hardware for Irregular, Low-Locality Applications

### 2012-07-23 doc/asplos13/intro.tex
moved latency paragraph

In this paper we introduce Grappa, a software runtime system that allows
a commodity x86 distributed-memory HPC cluster to be programmed as if it
were a single large shared-memory machine and provides scalable
performance for irregular applications. Grappa is designed to smooth
over some of the performance discontinuities in commodity hardware,
giving good performance when there is little locality to be exploited
while allowing the programmer to exploit it when it is available. Grappa
leverages as much freely available and commodity infrastructure as
possible. We use unmodified Linux for the operating system. An off the
shelf user-mode infiniband device driver stack~\cite{OFED} is used
for network interfacing. MPI is used for process setup and
tear down. GASnet~\cite{gasnet:spec} is used as the underlying mechanism for
remote memory reads and writes using active message invocations. To this
commodity hardware and software mix Grappa adds three main software
components: (1) a \emph{lightweight tasking} layer that supports a
context switch in as few as 38ns and distributed load balancing; (2) a
\emph{distributed shared memory} layer that supports customary
operations such as \emph{read and write as well as remote operations}
such as fetch-and-add~\cite{fetchandadd}; and (3) a \emph{message
aggregation} layer that combines short messages to mitigate the
aforementioned problem that commodity networks are designed to achieve
peak bandwidth only on large packet sizes, yet irregular applications
tend to fetch only a handful of bytes at a time.

>>>>

In this paper we introduce Grappa, a software runtime system that allows
a commodity x86 distributed-memory HPC cluster to be programmed as if it
were a single large shared-memory machine and provides scalable
performance for irregular applications. Grappa is designed to smooth
over some of the performance discontinuities in commodity hardware,
giving good performance when there is little locality to be exploited
while allowing the programmer to exploit it when it is available.

Grappa leverages as much freely available and commodity infrastructure as
possible. We use unmodified Linux for the operating system and an
off-the-shelf user-mode infiniband device driver stack~\cite{OFED}. MPI is
used for process setup and tear down. GASnet~\cite{gasnet:spec} is used as the
underlying mechanism for remote memory reads and writes using active message
invocations. To this commodity hardware and software mix Grappa adds three
main software components: (1) a \emph{lightweight tasking} layer that supports
a context switch in as few as 38ns and distributed global load balancing; (2)
a \emph{distributed shared memory} layer that supports customary operations
such as \emph{read and write as well as remote operations} such as
fetch-and-add~\cite{fetchandadd}; and (3) a \emph{message aggregation} layer
that combines short messages to mitigate the aforementioned problem that
commodity networks are designed to achieve peak bandwidth only on large packet
sizes, yet irregular applications tend to fetch only a handful of bytes at a
time. As we will show later, Grappa can tolerate latencies way beyond that of
the network. Therefore, Grappa can afford to \emph{trade latency for
throughput}: by {\em increasing} latency in key components of the system we
are able to increase our aggregate random access bandwidth (by delaying and
aggregating messages), our synchronization bandwidth (by delegating operations
to remote nodes), and our ability to improve load imbalance (work stealing
increases latency).
